---
title: "Jobs Report Development Notebook"
date: "Published on `r format(Sys.time(), '%m-%d-%y')`"
format:
  html:
    toc: true
    toc-location: left
    toc-depth: 5
    page-layout: full
fig-cap-location: top
editor:
  markdown:
    wrap: 72
execute: 
  echo: true
  output: true
---

# Load Packages to use 

To input data from ipums, we're going to use the ipumsr package

For packages that aren't yet installed:
```{r}
#install.packages("ipumsr")
```

Load packages
```{r}
#| output: false
library("tidyverse")
library("merced.clc")
library("ipumsr")
```

The ipums API allows us to talk directly to the ipums data source using an API key. More 
information [here](https://developer.ipums.org/docs/v2/get-started/).

Separately, there is an R package called `ipumsr` [here](https://tech.popdata.org/ipumsr/reference/set_ipums_api_key.html), which allows us to
talk to the IPUMS api directly via R code in Rstudio.

The ipumsr package requires an API to be set. Once an API key has been created for your 
IPUMS account, you want to save it to your local `.Renviron` file.

:::{.callout-warning}
LUKE NEEDS TO FIGURE OUT SYS.SETENV LATER
:::

First, we'll save our api key to the .Renviron like this.
```{r}
# Sys.setenv(ipums_api_key = "alasdf;lkjasdfl;kj2q12l;kjas;lf")
```

Then we tell the ipumsr package to use that key when talking to the ipums API data source. You should
only have to run this command the first time you set up.
```{r}
# ipumsr:::set_ipums_api_key(Sys.getenv("ipums_api_key"),
#                            save = TRUE)
```


# Downloading ACS Data

To download ACS data, you need to:

1. Specify a data collection (e.g. "CPS")

2. Specify a Description for your data download (e.g. "Jobs report feb 2020")

3. Specify the variables within the dataset that you want to download (e.g. "OCC" for Occupation code)

4. Specify a data sample (e.g. "cps2025_12s" for the december 2025 ipums cps data)

## Specify a data collection
```{r}
my_data_collection = "cps"
```

## Specify a Description
```{r}
my_data_description = "Jobs report feb 2020"
```

## Specify the variables within the dataset
```{r}
#| output: true
my_ipums_vars <- tribble(~`IPUMS code`, ~Description,
        "CITIZEN", "CITIZEN reports the citizenship status of foreign-born persons",
        "RACE", "Racial categories in the CPS")

my_ipums_vars %>% 
  arrange(`IPUMS code`) %>% 
  merced.clc::dt_table(filter = F,
                       search = F,
                       row_nums = F,
                       page = T)
```

## Specify a data sample
```{r}
my_data_samples = "cps2025_12s"
```


## Making an IPUMS API Data Request

### Creating an API request

Use the `ipumsr::define_extract_micro` function to define the data request that we're 
going to send to the IPUMS API.
```{r, download ipums extract}
#| eval: false
ipums_request <- ipumsr::define_extract_micro(collection = my_data_collection,
                                              description = my_data_description,
                                              variables = my_ipums_vars$`IPUMS code`,
                                              samples = my_data_samples)
```


### Submitting an API Request
```{r}
#| eval: false
ipums_extract <- ipumsr::submit_extract(ipums_request)
```


### Downloading an API Request
```{r}
#| eval: false
downloadable_extract <- ipumsr::wait_for_extract(ipums_extract)
path_to_data_files <- ipumsr::download_extract(downloadable_extract,
                                               download_dir = here::here("data"))
```








